---
title: "Data Visualization"
format: html
description: "This is a visualization of my personal data set collected Winter Quarter 2026 for Environmental Studies 193DS."
author: 
- name: "Sequoia Marx"
date: "03-10-2026"
---


## Reading in all of the necessary packages for these visualizations

```{r}
#| label: Set Up
#| message: false
#| echo: true
#| code-fold: true
library(tidyverse) #loading in tidyverse
library(here) #loading in here
library(janitor) #loading in janitor
library(readxl) #loading in readxl
library(ggeffects) #loading in ggeffects
library(knitr) #loading in knitr
my_data <- read_csv("my_data Sheet2 copy.csv"
  )
```

## Creating a Jitter Plot to display the times I woke up for school and non-school days, the median, and the interquartile ranges

```{r}
#| label: Jitter Plot displaying wake up time for school and non school days
#| echo: true
#| code-fold: true
my_data_clean <- my_data |> #labeling new data frame
clean_names() #cleaned names to only include lower case letters and underscores
ggplot(data = my_data_clean, #using my_data_clean data frame
aes(x = school_day_yes_no, #placing school_day_yes_no on the x-axis
y = wake_up_time_hrs_min, #placing wake_up_time_hrs_min on the y-axis
color = school_day_yes_no)) + #color by school_day_yes_no
geom_boxplot() + #creating a boxplot
labs(x = "School Day", #labeling the x-axis as School Day
y = "Wake Up time (hrs:min)", #labeling the y-axis as Wake Up time (hrs:min)
title = "Wake Up Time for School Days vs Non-School Days", #creating title
subtitle = "03/01/2026") + #creating subtitle
geom_jitter() + #creating jittered points
theme_light() + #changing theme to light
scale_color_manual(values = c("yes" = "blue4",
"no" = "red4")) + #changing colors based on whether or not it is a school day
theme(legend.position = "none") #removing the legend
```

**Figure 1.** This figure shows the distribution of recorded wake up times between school days, which is denoted by the blue points, and non-school days, which is denoted by the red points. Additionally, the figure shows the medians and quartiles for each type of day, represented by the boxes and lines, indicating that the median wake up time is later on non-school days than on school days.
